Parsing Postfix Log Files

### 2009-04-22 doc/logparser-glossary.tex
Corrections after proofreading chapters 6 & 7 and the appendices.

\newglossaryentry{Backscatter}{name={Backscatter},description={
    When a spam sender or worm sends mail with forged sender addresses,
    innocent sites are flooded with undeliverable mail notifications; this
    is called backscatter mail.
}}

\newglossaryentry{Joe-job}{name={Joe~Job},description={
    A joe~job describes spam mail sent using a faked sender address with
    the intention of sullying the good name of the user of that address.
    Joe~jobs are one cause of backscatter\glsadd{Backscatter}.
}}

@@

\newglossaryentry{<>}{name={$<>$},sort={<>},description={
    $<>$ is the sender address used for mail that should not be replied to,
    e.g.\ bounce notifications.  In the \acronym{SMTP} dialog all addresses
    are enclosed in $<>$, so \texttt{username@domain} is sent as
    \texttt{$<$username@domain$>$}; thus $<>$ is actually an empty address,
    but is always written as $<>$ for clarity.  All mail servers must
    accept mail sent from $<>$, or they are in violation of
    \acronym{RFC}~2821~\cite{RFC2821}.
}}

\newglossaryentry{queueid}{name={queueid},description={
    Each mail in Postfix's queue is assigned a queueid to uniquely identify
    it.  Queueids are assigned from a limited pool, so although they are
    guaranteed to be unique for the lifetime of the mail, they may be
    reused later.
}}

\newglossaryentry{hash}{name={hash},description={
    A hashing function transforms a string of characters to a number.  A
    common usage is to maintain a data structure indexed by strings in an
    efficient manner.  A full description is beyond the scope of this
    paper, further information can be found at~\cite{hash-functions}.
}}

\newglossaryentry{awk}{name={AWK},description={
    AWK is a general purpose programming language that is designed for
    processing text, and is available as a standard utility on all Unix
    systems.
}}

>>>>

\newglossaryentry{Backscatter}{name={Backscatter},description={
    When a spam sender or virus sends mail with forged sender addresses,
    innocent servers are flooded with undeliverable mail notifications;
    this is called backscatter.
}}

\newglossaryentry{Joe-job}{name={Joe~Job},description={
    A joe~job describes a large amount of spam mail sent using a faked
    sender address with the intention of sullying the good name of the user
    of that address.  Joe~jobs are one cause of
    backscatter\glsadd{Backscatter}.
}}

@@

\newglossaryentry{<>}{name={$<>$},sort={<>},description={
    $<>$ is the sender address used for sending bounce notifications.  In
    the \acronym{SMTP} conversation all addresses are enclosed in $<>$, so
    \texttt{username@domain} is sent as \texttt{$<$username@domain$>$};
    thus $<>$ is actually an empty address, but it is always written as
    $<>$ for clarity.  All mail servers must accept mail sent from $<>$, or
    they are in violation of \acronym{RFC}~2821~\cite{RFC2821}.
}}

\newglossaryentry{queueid}{name={queueid},description={
    Each mail in Postfix's queue is assigned a queueid to uniquely identify
    it.  Queueids are assigned from a limited pool, so although they are
    guaranteed to be unique for the lifetime of the mail they are assigned
    to, they may be reused later.
}}

\newglossaryentry{hash}{name={hash},description={
    A hashing function transforms a string of characters to a number.  A
    common usage is to maintain a data structure indexed by strings in an
    efficient manner.  A full description is beyond the scope of this
    thesis, further information can be found at~\cite{hash-functions}.
}}

\newglossaryentry{awk}{name={AWK},description={
    AWK is a general purpose programming language that is designed for
    processing text and is available as a standard utility on all Unix
    systems.
}}

### 2009-04-22 doc/logparser-other-parsers-review.tex
Corrections after proofreading chapters 6 & 7 and the appendices.

Splunk aims to index all of an organisation's log files, providing a
centralised view capable of searching and correlating diverse log sources.
The web interface provides search functionality, generating statistics and
graphs in real time, a facility not provided by \parsername{}.  Splunk
allows quite complicated searches, based on the fields it extracts
(described later) or the full text of the log line, though it is not
possible to search on partial words.  Search results are not available for
use with other tools.  Searches can be saved for reuse; saved searches can
be run periodically and the results mailed to a recipient or sent to an
external program for further processing.  The author was unable to save
searches, though that may have been because of limitations in the free
version.  The web interface is optimised for interactive use rather than
automated queries, and it does not appear to be possible to write
independent tools to utilise the Splunk database, whereas \parsername{}
provides the database and leaves it to the user to utilise it without limit
or restriction.  Some additional Postfix reports are supposedly available
on \urlLastChecked{http://www.splunkbase.com/}{2008/11/23}, but the author
was unable to find any Postfix reports, or indeed reports for any other log
file types: every category was empty, even those that the web site claimed
had numerous reports available.  Many types of graphs can be generated,
though all except the bubble and heatmap graphs are variations of a bar or
pie chart.  It is trivial to drill down through the graphs to select a
portion of the data, e.g.\ select the hour with the largest number of
events, then select a particular host, and finally a specific sender
address.  All searches performed using the indexed data returned reasonable
results.  The full power of \acronym{SQL} is available when searching the
data extracted by \parsername{}, allowing the user to search on arbitrarily
complicated conditions.

@@

Copious documentation is made available on
\urlLastChecked{http://www.splunk.com/}{2008/11/23}, but the abundance of
material and lack of organisation makes it hard to find the topic being
sought, and searches confusingly tended to return results from old
documentation rather than new.  In general the documentation appears to
have been written by someone intimately acquainted with the software, who
has difficulty understanding how a newcomer would approach tasks or the
questions they would ask.

@@

AWStats can produce simple graphs from many different services' log files,
but supporting numerous log file formats without special purpose code
limits its functionality.  The data it can extract from Postfix log files
is limited in comparison to \parsername{}: time2, email, email\_r, host,
host\_r, method, url, code, and bytesd.  No explanation for any of those
fields is provided in the documentation at
\urlLastChecked{http://awstats.sourceforge.net/docs/awstats_faq.html\#MAIL}{2008/11/23},
so the author could neither understand what data is extracted, nor
determine what data is missing in comparison to \parsername{}, which fully
documents all the data it extracts.  AWStats coerces Postfix log files into
the log file format used by the Apache web server, for analysis by AWStats'
HTTP log file parser.  The converting parser only deals with a small
portion of the log lines generated by Postfix, silently skipping those it
cannot parse, and does not distinguish between different delivery attempt
rejections; extending it to parse all log lines would be at least as much
work as writing a new parser.  It does correlate log lines by queueid (not
by pid), but it does not deal with any of the other complications described
in \sectionref{complications}.  AWStats supports saving data extracted from
log files, but the format of the data store is not documented.  It also
supports reading compressed log files, but that was not tested.

@@

YAALA provides a mini-language based on \acronym{SQL} that is used when
generating reports; sample reports can be seen
at~\urlLastChecked{http://www.yaala.org/samples.html}{2008/11/23}.  Example
query for HTTP proxy servers: \newline{} \tab{} \texttt{requests BY file
WHERE host =\~{} Google} \newline{} The mini-language is quite limited and
cannot be used to extract data for external use, merely to create reports.
Only data selected by the query will be saved in the data store; other data
will be discarded, and removed from the data store if already present.

>>>>

Splunk aims to index all of an organisation's log files, providing a
centralised view capable of searching and correlating diverse log sources.
The web interface provides search functionality, generating statistics and
graphs in real time, a facility not provided by \parsername{}.  Splunk
allows quite complicated searches, based on the fields it extracts
(described later) or the full text of the log line, though it is not
possible to search on partial words.  Search results are not available for
use with other tools.  Searches can be saved for reuse; saved searches can
be run periodically and the results mailed to a recipient or sent to an
external program for further processing.  The author was unable to save
searches, though that may have been because of limitations in the free
version.  The web interface is optimised for interactive use rather than
automated queries, and it does not appear to be possible to write
independent tools to utilise the Splunk database, whereas \parsername{}
provides the database and leaves it to the user to utilise it without limit
or restriction.  Some additional Postfix reports are supposedly available
on \urlLastChecked{http://www.splunkbase.com/}{2009/04/21}, but the author
was unable to find any Postfix reports, or indeed reports for any other log
file types: every category was empty, even those that the web site claimed
had numerous reports available.  Many types of graphs can be generated,
though all except the bubble and heatmap graphs are variations of a bar or
pie chart.  It is trivial to drill down through the graphs to select a
portion of the data, e.g.\ select the hour with the largest number of
events, then select a particular host, and finally a specific sender
address.  All searches performed using the indexed data returned reasonable
results.  The full power of \acronym{SQL} is available when searching the
data extracted by \parsername{}, allowing the user to search on arbitrarily
complicated conditions.

@@

Copious documentation is made available on
\urlLastChecked{http://www.splunk.com/}{2009/04/21}, but the abundance of
material and lack of organisation makes it hard to find the topic being
sought, and searches confusingly tended to return results from old
documentation rather than new.  In general the documentation appears to
have been written by someone intimately acquainted with the software, who
has difficulty understanding how a newcomer would approach tasks or the
questions they would ask.

@@

AWStats can produce simple graphs from many different services' log files,
but supporting numerous log file formats without special purpose code
limits its functionality.  The data it can extract from Postfix log files
is limited in comparison to \parsername{}: time2, email, email\_r, host,
host\_r, method, url, code, and bytesd.  No explanation for any of those
fields is provided in the documentation at
\urlLastChecked{http://awstats.sourceforge.net/docs/awstats_faq.html\#MAIL}{2009/04/21},
so the author could neither understand what data is extracted, nor
determine what data is missing in comparison to \parsername{}, which fully
documents all the data it extracts.  AWStats coerces Postfix log files into
the log file format used by the Apache web server, for analysis by AWStats'
HTTP log file parser.  The converting parser only deals with a small
portion of the log lines generated by Postfix, silently skipping those it
cannot parse, and does not distinguish between different delivery attempt
rejections; extending it to parse all log lines would be at least as much
work as writing a new parser.  It does correlate log lines by queueid (not
by pid), but it does not deal with any of the other complications described
in \sectionref{complications}.  AWStats supports saving data extracted from
log files, but the format of the data store is not documented.  It also
supports reading compressed log files, but that was not tested.

@@

YAALA provides a mini-language based on \acronym{SQL} that is used when
generating reports; sample reports can be seen at
\urlLastChecked{http://www.yaala.org/samples.html}{2009/04/21}.  Example
query for HTTP proxy servers: \newline{} \tab{} \texttt{requests BY file
WHERE host =\~{} Google} \newline{} The mini-language is quite limited and
cannot be used to extract data for external use, merely to create reports.
Only data selected by the query will be saved in the data store; other data
will be discarded, and removed from the data store if already present.

### 2009-04-22 doc/logparser-postfix-daemons.tex
Corrections from proofreading for chapters 4 and 5.
Corrections after proofreading chapters 6 & 7 and the appendices.

% vim: set filetype=tex :
% Postfix components
\newglossaryentry{bounce}{name={bounce},type={postfix},description={
    The bounce daemon is responsible for sending bounce notifications in
    Postfix version 2.3 and later.
    \daemonDocURL{http://www.postfix.org/bounce.8.html}{2009/02/23}.
}}

@@

\newglossaryentry{local}{name={local},type={postfix},description={
    Local is the Postfix component responsible for local delivery of mail
    (i.e.\ mail delivered on the server Postfix is running on); this
    includes alias expansion, processing of a user's \texttt{.forward}
    file, and delivery of the mail, whether to a user's mailbox or a
    program such as a mailing list manager.
    \daemonDocURL{http://www.postfix.org/local.8.html}{2009/02/23}.
}}

\newglossaryentry{pickup}{name={pickup},type={postfix},description={
    Pickup is the daemon that deals with mail submitted locally via
    \daemon{postdrop} and \daemon{sendmail,} passing the mail on to
    \daemon{cleanup} for further processing.
    \daemonDocURL{http://www.postfix.org/pickup.8.html}{2009/02/23}.
}}

\newglossaryentry{postdrop}{name={postdrop},type={postfix},description={
    Postdrop is used when submitting mail locally on the server: it copies
    its input into a newly created mail in the queue, for processing by
    \daemon{pickup} and subsequent delivery by one or more of
    \daemon{lmtp}, \daemon{local}, \daemon{smtp}, or \daemon{virtual}.
    \daemonDocURL[\newline{}]{http://www.postfix.org/postdrop.1.html}{2009/02/23}.
}}

@@

\newglossaryentry{virtual}{name={virtual},type={postfix},description={
    The Postfix component responsible for delivery of mails to virtual
    domains.  When \daemon{local} delivers mail the destination is
    determined only by the portion of the email address on the left side of
    the \texttt{@}, whereas with \daemon{virtual} delivery the destination
    is determined by the entire email address.  E.g.\ if the server is
    responsible for both the \texttt{example.org} and \texttt{example.net}
    domains: \daemon{local} would deliver mail for
    \texttt{john@example.org} and \texttt{john@example.net} to the same
    mailbox, whereas \daemon{virtual} would deliver mail for those
    addresses to different mailboxes.  Virtual delivery is used where the
    local part of an address may be present in multiple domains, and each
    must be delivered to different users.
    \daemonDocURL{http://www.postfix.org/virtual.8.html}{2009/02/23}.
}}

>>>>

% vim: set filetype=tex :
% Postfix components
\newglossaryentry{bounce}{name={bounce},type={postfix},description={
    The bounce daemon is responsible for generating bounce notifications in
    Postfix version 2.3 and later.
    \daemonDocURL{http://www.postfix.org/bounce.8.html}{2009/02/23}.
}}

@@

\newglossaryentry{local}{name={local},type={postfix},description={
    The Postfix component responsible for local delivery of mail (i.e.\
    mail delivered on the server Postfix is running on); this includes
    alias expansion, processing of a user's \texttt{.forward} file, and
    delivery of the mail, whether to a user's mailbox or a program such as
    a mailing list manager.
    \daemonDocURL{http://www.postfix.org/local.8.html}{2009/02/23}.
}}

\newglossaryentry{pickup}{name={pickup},type={postfix},description={
    Pickup is the daemon that deals with mail submitted locally via
    \daemon{postdrop}, passing the mail on to \daemon{cleanup} for further
    processing.
    \daemonDocURL{http://www.postfix.org/pickup.8.html}{2009/02/23}.
}}

\newglossaryentry{postdrop}{name={postdrop},type={postfix},description={
    Postdrop is used when submitting mail locally on the server: it copies
    its input into a newly created mail in the queue, for processing by
    \daemon{pickup} and subsequent delivery.
    \daemonDocURL[\newline{}]{http://www.postfix.org/postdrop.1.html}{2009/02/23}.
}}

@@

\newglossaryentry{virtual}{name={virtual},type={postfix},description={
    The Postfix component responsible for delivery of mails to virtual
    domains.  When \daemon{local} delivers mail, the destination is
    determined only by the portion of the email address on the left side of
    the \texttt{@}, whereas when \daemon{virtual} delivers mail, the
    destination is determined by the entire email address.  For example, if
    the server is responsible for both the \texttt{example.org} and
    \texttt{example.net} domains: \daemon{local} would deliver mail for
    \texttt{john@example.org} and \texttt{john@example.net} to the same
    mailbox, whereas \daemon{virtual} would deliver mail for those
    addresses to different mailboxes.  Virtual delivery is used where the
    local part of an address may be present in multiple domains, and each
    must be delivered to different users.
    \daemonDocURL{http://www.postfix.org/virtual.8.html}{2009/02/23}.
}}
